---
title:    "Ocampo Lab mouse reprogramming"
subtitle: "Selected genes expression"
author:   "Author: João Agostinho de Sousa"
date:     "Date: `r format(Sys.time(),'%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: TRUE
    toc_float: TRUE
    theme: default
    code_folding: hide
editor_options:
  chunk_output_type: console
---

# Load libraries
```{r}

library(tidyverse)
library(SummarizedExperiment)
library(patchwork)
library(ComplexHeatmap)

```

# Load objects
```{r}

# SummarizedExperiment object
load(file = "output/database/SummarizedExperiment_object.RData")

# Log2 normalized counts object
load(file = "output/database/log2_norm_counts_object.RData")

# DGE results
load(file = "output/database/DGE_results.RData")

# Selected genes
load(file = "output/database/selected_genes.RData")

```

# Load functions
```{r}

# Theme publication
source(file = "analysis/functions/theme_publication.R")

```

# Create data for the heatmaps
```{r}

# Heatmap plot data
heatmap_plot_data <- 
  map(selected_genes,
  ~ log2_norm_counts_object$log2_norm_counts_long %>% 
  filter(gene_name %in% .x) %>% 
  mutate(gene_name = case_when(is.na(gene_name) ~ gene_id,
                               TRUE             ~ gene_name)) %>% 
  group_by(Genotype, Tissue, gene_id, gene_name) %>%
  summarise(mean_log2_norm_counts = mean(log2_norm_counts, na.rm = T)) %>%
  ungroup
  )

```

# Heatmap colors
```{r}

heatmap_colors <- circlize::colorRamp2(seq(0, 15, 16/11), c(rev(RColorBrewer::brewer.pal(n = 11, name = "RdBu"))))
heatmap_colors_scaled <- circlize::colorRamp2(seq(-2, 2, 4/10), c(rev(RColorBrewer::brewer.pal(n = 11, name = "RdBu"))))

```

# 1. Hepatocyte genes on the liver data
```{r}

########################################
# Non-scaled
########################################
ht1 <- 
  heatmap_plot_data$Hepatocyte_Hishida %>% 
  filter(Tissue == "Liver") %>%
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  ComplexHeatmap::Heatmap(name                 = "Log2(norm. counts + 1)",
                          col                  = heatmap_colors, 
                          cluster_columns      = F, 
                          column_title         = "Liver",
                          column_title_gp      = gpar(fontface = "bold", cex = 1.25),
                          row_title            = "Hepatocyte",
                          row_title_rot        = 0,
                          row_title_side       = "right",
                          row_title_gp         = gpar(fontface = "bold"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")

########################################
# Scaled
########################################
ht1_scaled <- 
  heatmap_plot_data$Hepatocyte_Hishida %>% 
  filter(Tissue == "Liver") %>%
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  t %>% 
  scale %>% 
  t %>% 
  ComplexHeatmap::Heatmap(name                 = "Row Z-score",
                          col                  = heatmap_colors_scaled, 
                          cluster_columns      = F, 
                          column_title         = "Liver",
                          column_title_gp      = gpar(fontface = "bold", cex = 1.25),
                          row_title            = "Hepatocyte",
                          row_title_rot        = 0,
                          row_title_side       = "right",
                          row_title_gp         = gpar(fontface = "bold"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")

```

# 2. Hepatocyte genes on the liver data
```{r}

########################################
# Non-scaled
########################################
ht2 <- 
  rbind(heatmap_plot_data$Early_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Early"),
        heatmap_plot_data$Early_and_advanced_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Early and advanced")) %>% 
  rbind(heatmap_plot_data$Advanced_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Advanced")) %>% 
  filter(Tissue == "Small intestine") %>%
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts, gene_category) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  mutate(gene_category = factor(gene_category, levels = c("Early", "Early and advanced", "Advanced"))) %>% 
  arrange(gene_category) %>% 
  dplyr::select(-gene_category) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  ComplexHeatmap::Heatmap(name                 = "Log2(norm. counts + 1)",
                          col                  = heatmap_colors, 
                          split                = c(rep(1, length(selected_genes$Early_intermediate_reprogramming_Chondronasiou)),
                                                   rep(2, length(selected_genes$Early_and_advanced_intermediate_reprogramming_Chondronasiou)),
                                                   rep(3, length(selected_genes$Advanced_intermediate_reprogramming_Chondronasiou))),
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          column_title         = "Small intestine",
                          column_title_gp      = gpar(fontface = "bold", cex = 1.25),
                          row_title            = c("Early", "Early and advanced", "Advanced"),
                          row_title_rot        = 0,
                          gap                  = unit(.1, "in"),
                          row_title_side       = "right",
                          row_title_gp         = gpar(fontface = "bold"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")

########################################
# Scaled
########################################
ht2_scaled <- 
  rbind(heatmap_plot_data$Early_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Early"),
        heatmap_plot_data$Early_and_advanced_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Early and advanced")) %>% 
  rbind(heatmap_plot_data$Advanced_intermediate_reprogramming_Chondronasiou %>% mutate(gene_category = "Advanced")) %>% 
  filter(Tissue == "Small intestine") %>%
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts, gene_category) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  mutate(gene_category = factor(gene_category, levels = c("Early", "Early and advanced", "Advanced"))) %>% 
  arrange(gene_category) %>% 
  dplyr::select(-gene_category) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  t %>% 
  scale %>% 
  t %>% 
  ComplexHeatmap::Heatmap(name                 = "Row Z-score",
                          col                  = heatmap_colors_scaled, 
                          split                = c(rep(1, length(selected_genes$Early_intermediate_reprogramming_Chondronasiou)),
                                                   rep(2, length(selected_genes$Early_and_advanced_intermediate_reprogramming_Chondronasiou)),
                                                   rep(3, length(selected_genes$Advanced_intermediate_reprogramming_Chondronasiou))),
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          column_title         = "Small intestine",
                          column_title_gp      = gpar(fontface = "bold", cex = 1.25),
                          row_title            = c("Early", "Early and advanced", "Advanced"),
                          row_title_rot        = 0,
                          gap                  = unit(.1, "in"),
                          row_title_side       = "right",
                          row_title_gp         = gpar(fontface = "bold"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")

```

# 3. Selected genes on data per tissue
```{r}

########################################
# Non-Scaled
########################################
ht3 <- 
map(list(
  "Progenitor"            = heatmap_plot_data$Progenitor_Hishida,
  "p53"                   = heatmap_plot_data$p53_Hishida,
  "Cell_cycle"            = heatmap_plot_data$Cell_cycle_Hishida,
  "Epigenetic_regulators" = heatmap_plot_data$Epigenetic_regulators_Hishida,
  "Apoptosis"             = heatmap_plot_data$Apoptosis), function(x){
    
map(levels(SummarizedExperiment_object$Tissue),
  
  ~ x %>% 
  filter(Tissue == .x) %>% 
  dplyr::select(-Tissue) %>% 
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  ComplexHeatmap::Heatmap(name                 = "Log2(norm. counts + 1)",
                          col                  = heatmap_colors,
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          column_title         = .x,
                          row_title_rot        = 0,
                          column_title_gp      = gpar(fontface = "bold", cex = 1),
                          gap                  = unit(.1, "in"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")
  )
})

########################################
# Scaled
########################################
ht3_scaled <- 
map(list(
  "Progenitor"            = heatmap_plot_data$Progenitor_Hishida,
  "p53"                   = heatmap_plot_data$p53_Hishida,
  "Cell_cycle"            = heatmap_plot_data$Cell_cycle_Hishida,
  "Epigenetic_regulators" = heatmap_plot_data$Epigenetic_regulators_Hishida,
  "Apoptosis"             = heatmap_plot_data$Apoptosis), function(x){
    
map(levels(SummarizedExperiment_object$Tissue),
  
  ~ x %>% 
  filter(Tissue == .x) %>% 
  dplyr::select(-Tissue) %>% 
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  t %>% 
  scale %>% 
  t %>% 
  ComplexHeatmap::Heatmap(name                 = "Row Z-score",
                          col                  = heatmap_colors_scaled,
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          column_title         = .x,
                          row_title_rot        = 0,
                          column_title_gp      = gpar(fontface = "bold", cex = 1),
                          gap                  = unit(.1, "in"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")
  )
})

```

# 4. Selected genes on data with combined tissues
```{r}

# Heatmap plot data
heatmap_plot_combined_data <- 
  map(selected_genes,
  ~ log2_norm_counts_object$log2_norm_counts_long %>% 
  filter(gene_name %in% .x) %>% 
  mutate(gene_name = case_when(is.na(gene_name) ~ gene_id,
                               TRUE             ~ gene_name)) %>% 
  group_by(Genotype, gene_id, gene_name) %>%
  summarise(mean_log2_norm_counts = mean(log2_norm_counts, na.rm = T)) %>%
  ungroup
  )

########################################
# Non-Scaled
########################################
ht4 <- 
map(list(
  "Progenitor"            = heatmap_plot_combined_data$Progenitor_Hishida,
  "p53"                   = heatmap_plot_combined_data$p53_Hishida,
  "Cell_cycle"            = heatmap_plot_combined_data$Cell_cycle_Hishida,
  "Epigenetic_regulators" = heatmap_plot_combined_data$Epigenetic_regulators_Hishida,
  "Apoptosis"             = heatmap_plot_combined_data$Apoptosis), function(x){
    
  x %>% 
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>% 
  ComplexHeatmap::Heatmap(name                 = "Log2(norm. counts + 1)",
                          col                  = heatmap_colors,
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          row_title_rot        = 0,
                          column_title_gp      = gpar(fontface = "bold", cex = 1),
                          gap                  = unit(.1, "in"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")
})

########################################
# Scaled
########################################
ht4_scaled <- 
map(list(
  "Progenitor"            = heatmap_plot_combined_data$Progenitor_Hishida,
  "p53"                   = heatmap_plot_combined_data$p53_Hishida,
  "Cell_cycle"            = heatmap_plot_combined_data$Cell_cycle_Hishida,
  "Epigenetic_regulators" = heatmap_plot_combined_data$Epigenetic_regulators_Hishida,
  "Apoptosis"             = heatmap_plot_combined_data$Apoptosis), function(x){
    
  x %>% 
  dplyr::select(gene_name, Genotype, mean_log2_norm_counts) %>% 
  spread(Genotype, mean_log2_norm_counts) %>% 
  column_to_rownames("gene_name") %>% 
  as.matrix() %>%
  t %>% 
  scale %>% 
  t %>% 
  ComplexHeatmap::Heatmap(name                 = "Row Z-score",
                          col                  = heatmap_colors_scaled,
                          cluster_columns      = F, 
                          cluster_rows         = T, 
                          cluster_row_slices   = F,
                          row_title_rot        = 0,
                          column_title_gp      = gpar(fontface = "bold", cex = 1),
                          gap                  = unit(.1, "in"),
                          column_names_rot     = 45,
                          rect_gp              = gpar(col = "white", lwd = 1),
                          heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(4, "cm")),
                          border               = "black")
})

```

# Save heatmaps 1
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht1.pdf", width = 4, height = 7)
draw(ht1, heatmap_legend_side = "bottom")
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht1_scaled.pdf", width = 4, height = 7)
draw(ht1_scaled, heatmap_legend_side = "bottom")
dev.off()

```

# Save heatmaps 2
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht2.pdf", width = 4.5, height = 7)
draw(ht2, heatmap_legend_side = "bottom")
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht2_scaled.pdf", width = 4.5, height = 7)
draw(ht2_scaled, heatmap_legend_side = "bottom")
dev.off()

```

# Save heatmaps 3
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht3_Progenitor.pdf", width = 11, height = 4)
(ht3$Progenitor[[1]] + 
 ht3$Progenitor[[2]] + 
 ht3$Progenitor[[3]] + 
 ht3$Progenitor[[4]] + 
 ht3$Progenitor[[5]] + 
 ht3$Progenitor[[6]] + 
 ht3$Progenitor[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Progenitor"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_p53.pdf", width = 11, height = 2.5)
(ht3$p53[[1]] + 
 ht3$p53[[2]] + 
 ht3$p53[[3]] + 
 ht3$p53[[4]] + 
 ht3$p53[[5]] + 
 ht3$p53[[6]] + 
 ht3$p53[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("p53"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Cell_cycle.pdf", width = 11, height = 4)
(ht3$Cell_cycle[[1]] + 
 ht3$Cell_cycle[[2]] + 
 ht3$Cell_cycle[[3]] + 
 ht3$Cell_cycle[[4]] + 
 ht3$Cell_cycle[[5]] + 
 ht3$Cell_cycle[[6]] + 
 ht3$Cell_cycle[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Cell cycle"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Epigenetic_regulators.pdf", width = 11, height = 8)
(ht3$Epigenetic_regulators[[1]] + 
 ht3$Epigenetic_regulators[[2]] + 
 ht3$Epigenetic_regulators[[3]] + 
 ht3$Epigenetic_regulators[[4]] + 
 ht3$Epigenetic_regulators[[5]] + 
 ht3$Epigenetic_regulators[[6]] + 
 ht3$Epigenetic_regulators[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Epigenetic regulators"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Apoptosis.pdf", width = 11, height = 11)
(ht3$Apoptosis[[1]] + 
 ht3$Apoptosis[[2]] + 
 ht3$Apoptosis[[3]] + 
 ht3$Apoptosis[[4]] + 
 ht3$Apoptosis[[5]] + 
 ht3$Apoptosis[[6]] + 
 ht3$Apoptosis[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Apoptosis"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

```

# Save heatmaps 3 scaled
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht3_Progenitor_scaled.pdf", width = 11, height = 4)
(ht3_scaled$Progenitor[[1]] + 
 ht3_scaled$Progenitor[[2]] + 
 ht3_scaled$Progenitor[[3]] + 
 ht3_scaled$Progenitor[[4]] + 
 ht3_scaled$Progenitor[[5]] + 
 ht3_scaled$Progenitor[[6]] + 
 ht3_scaled$Progenitor[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Progenitor"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_p53_scaled.pdf", width = 11, height = 2.5)
(ht3_scaled$p53[[1]] + 
 ht3_scaled$p53[[2]] + 
 ht3_scaled$p53[[3]] + 
 ht3_scaled$p53[[4]] + 
 ht3_scaled$p53[[5]] + 
 ht3_scaled$p53[[6]] + 
 ht3_scaled$p53[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("p53"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Cell_cycle_scaled.pdf", width = 11, height = 4)
(ht3_scaled$Cell_cycle[[1]] + 
 ht3_scaled$Cell_cycle[[2]] + 
 ht3_scaled$Cell_cycle[[3]] + 
 ht3_scaled$Cell_cycle[[4]] + 
 ht3_scaled$Cell_cycle[[5]] + 
 ht3_scaled$Cell_cycle[[6]] + 
 ht3_scaled$Cell_cycle[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Cell cycle"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Epigenetic_regulators_scaled.pdf", width = 11, height = 8)
(ht3_scaled$Epigenetic_regulators[[1]] + 
 ht3_scaled$Epigenetic_regulators[[2]] + 
 ht3_scaled$Epigenetic_regulators[[3]] + 
 ht3_scaled$Epigenetic_regulators[[4]] + 
 ht3_scaled$Epigenetic_regulators[[5]] + 
 ht3_scaled$Epigenetic_regulators[[6]] + 
 ht3_scaled$Epigenetic_regulators[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Epigenetic regulators"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht3_Apoptosis_scaled.pdf", width = 11, height = 11)
(ht3_scaled$Apoptosis[[1]] + 
 ht3_scaled$Apoptosis[[2]] + 
 ht3_scaled$Apoptosis[[3]] + 
 ht3_scaled$Apoptosis[[4]] + 
 ht3_scaled$Apoptosis[[5]] + 
 ht3_scaled$Apoptosis[[6]] + 
 ht3_scaled$Apoptosis[[7]]) %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Apoptosis"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

```

# Save heatmaps 4
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht4_Progenitor.pdf", width = 2.75, height = 4)
ht4$Progenitor %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Progenitor"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_p53.pdf", width = 2.75, height = 2.5)
ht4$p53 %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("p53"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Cell_cycle.pdf", width = 2.75, height = 4)
ht4$Cell_cycle %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Cell cycle"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Epigenetic_regulators.pdf", width = 2.75, height = 8)
ht4$Epigenetic_regulators %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Epigenetic regulators"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Apoptosis.pdf", width = 2.75, height = 11)
ht4$Apoptosis %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Apoptosis"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

```

# Save heatmaps 4 scaled
```{r}

pdf(file = "output/figures/02_selected_genes_expression/ht4_Progenitor_scaled.pdf", width = 2.75, height = 4)
ht4_scaled$Progenitor %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Progenitor"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_p53_scaled.pdf", width = 2.75, height = 2.5)
ht4_scaled$p53 %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("p53"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Cell_cycle_scaled.pdf", width = 2.75, height = 4)
ht4_scaled$Cell_cycle %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Cell cycle"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Epigenetic_regulators_scaled.pdf", width = 2.75, height = 8)
ht4_scaled$Epigenetic_regulators %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Epigenetic regulators"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

pdf(file = "output/figures/02_selected_genes_expression/ht4_Apoptosis_scaled.pdf", width = 2.75, height = 11)
ht4_scaled$Apoptosis %>% 
draw(heatmap_legend_side = "bottom", 
     row_title           = c("Apoptosis"), 
     row_title_side      = "right",
     row_title_gp        = gpar(fontface = "bold")
     )
dev.off()

```

# Session information
```{r Session info, echo = FALSE, results = "markup"}

devtools::session_info()

```

This document was processed on: `r Sys.Date()`.
